---
title: "Assignment 1"
author: "Iryna Feuerstein"
date: "5th of March, 2017"
output: html_document
---

# Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity_data <- read.csv("activity.csv")
summary(activity_data)
```

# What is the mean total number of steps taken per day?

### Calculating the total number of steps taken per day

```{r}
steps_per_day <- tapply(activity_data$steps, activity_data$date, FUN = sum, na.rm=TRUE)
```

### Histogramm of the total number of steps taken each day

```{r}
hist(steps_per_day)
```

### Calculating the mean and median of the total number of steps taken per day

```{r}
mean(steps_per_day)
median(steps_per_day)
```

# What is the average daily activity pattern?

### Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 

```{r}
avg_steps_per_interval <- tapply(activity_data$steps, activity_data$interval, FUN = mean, na.rm=TRUE)
plot(avg_steps_per_interval, type = "l", xlab = "Intervals", ylab = "Avarage number of steps")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_index <- which.max(avg_steps_per_interval)
avg_steps_per_interval[max_index]
```

# Imputing missing values

### Total number of missing values in the dataset
```{r}
summary(is.na(activity_data))
sum(is.na(activity_data$steps))
```

### Filling in missing values in the dataset
```{r}
filled_activity <- data.frame(activity_data)
filled_activity$steps <- ifelse(is.na(filled_activity$steps), avg_steps_per_interval[as.character(filled_activity$interval)], filled_activity$steps)

```

### Histogram of the total number of steps taken each day after filling in the missing data.
```{r}
filled_steps_per_day <- tapply(filled_activity$steps, filled_activity$date, FUN = sum)
hist(filled_steps_per_day)
```

### Mean and median total number of steps taken per day after filling in the missing data.
```{r}
mean(filled_steps_per_day)
median(filled_steps_per_day)
```

# Are there differences in activity patterns between weekdays and weekends?










