---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# Loading and preprocessing the data

```{r}
unzip("activity.zip")
activity_data <- read.csv("activity.csv")
summary(activity_data)
```

# What is the mean total number of steps taken per day?

### Calculating the total number of steps taken per day

```{r}
steps_per_day <- tapply(activity_data$steps, activity_data$date, FUN = sum, na.rm=TRUE)
```

### Histogramm of the total number of steps taken each day

```{r}
hist(steps_per_day)
```

### Calculating the mean and median of the total number of steps taken per day

```{r}
mean(steps_per_day)
median(steps_per_day)
```

# What is the average daily activity pattern?

### Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 

```{r}
avg_steps_per_interval <- tapply(activity_data$steps, activity_data$interval, FUN = mean, na.rm=TRUE)
plot(avg_steps_per_interval, type = "l", xlab = "Intervals", ylab = "Avarage number of steps")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_index <- which.max(avg_steps_per_interval)
avg_steps_per_interval[max_index]
```

# Imputing missing values

### Total number of missing values in the dataset
```{r}
summary(is.na(activity_data))
missing_values_amount <- sum(is.na(activity_data$steps))
```
There are `r missing_values_amount` missing values in the dataset.

### Filling in missing values in the dataset
For filling in the missing values the mean for that 5-minute interval is used.
```{r}
filled_activity <- data.frame(activity_data)
filled_activity$steps <- ifelse(is.na(filled_activity$steps), avg_steps_per_interval[as.character(filled_activity$interval)], filled_activity$steps)

```

### Histogram of the total number of steps taken each day after filling in the missing data.
```{r}
filled_steps_per_day <- tapply(filled_activity$steps, filled_activity$date, FUN = sum)
hist(filled_steps_per_day)
```

### Mean and median total number of steps taken per day after filling in the missing data.
```{r}
mean(filled_steps_per_day)
median(filled_steps_per_day)
```

# Are there differences in activity patterns between weekdays and weekends?

### Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
filled_activity$weekday_type <- "weekday"
filled_activity$date <- as.Date(activity_data$date)
activity_weekdays <- weekdays(filled_activity$date)
filled_activity$weekday_type <- ifelse(activity_weekdays == "Samstag" | activity_weekdays == "Sonntag", "weekend", "weekday")
```

### A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days and weekend days.
```{r}
activity_on_weekday <- subset(filled_activity, filled_activity$weekday_type=="weekday")
activity_on_weekend <- subset(filled_activity, filled_activity$weekday_type=="weekend")
avg_steps_per_interval_on_weekday <- tapply(activity_on_weekday$steps, activity_on_weekday$interval, FUN = mean)
avg_steps_per_interval_on_weekend <- tapply(activity_on_weekend$steps, activity_on_weekend$interval, FUN = mean)
par(mfrow=c(2,1)) 
plot(avg_steps_per_interval_on_weekday, type = "l", xlab = "Intervals", ylab = "Avarage number of steps", main = "On weekdays")
plot(avg_steps_per_interval_on_weekend, type = "l", xlab = "Intervals", ylab = "Avarage number of steps", main = "On weekend")
```


